figures and calculations for global fisheries management

### 2019-01-12 documents/figs-and-calcs.Rmd
worked on calculations

@@

@@

@@

Based on this calculation, we find that up to xx\% of global catch comes from TURFs and up to XX \% comes from cooperatives. In total, based on best available data to date,  we estimate that somewhere between XXX\% and XXX\% of global wild fish catch comes from all forms of RBFM combined. The What kinds of fisheries make up this RBFM catch? RBFM catches come from XX\% of the countries reporting  fish catches in 2016xx.

@@

@@

What factors drive the adoption of various forms of RBFM? Anecdotal evidence suggests that large, industrial fisheries, often for mobile species are most likely to adopt some form of ITQ, while small-scale artisinal fisheries, often for benthic species, are most likely to adopt a TURF. To test this expectation, we built a dataset of XXX ITQ fisheries and XXX TURF fisheries, and matched them to the fishery classification in \cite{Upside}. This provided us with a dataset of XXX fisheries with the country, species or genus name, and some other ecological and economic characteristics. Conditional on adopting some form of RBFM, our goal was to estimate the likelihood of adopting an ITQ versus a TURF. To do so, we ran a logistic regression with ITQ on the left ha

@@

to all XX non-rbfm fisheries in the (however many fisheries are not catch shares in projection_data)

@@

@@

Using that approach, authors have estimated a benefit of global RBFM between \$XXX Billion and \$XXX Billion per year. see Costello et al. 2016

>>>>

http://www.fao.org/3/i9540en/I9540EN.pdf
"The total production included 80.0 million tonnes of food fish (USD 231.6 billion)"
- farmed fish has exceeded wild caught

@@

= 7%

@@

projection_data_line65 <- readRDS(here("data","projection_data.rds")) %>% 
  filter(Scenario == "Historic") %>%
  group_by(IdOrig) %>% 
  filter(Year == max(Year))%>%
  filter(Dbase == "FAO") %>%
  filter(IdLevel != "Neis")%>%
  filter(CatchShare == "1")

itq_catch <- sum(projection_data_line65$Catch)

percent_global <- (itq_catch/80000000)*100
percent_global
#6.995134

@@

Based on this calculation, we find that up to xx\% (1.7%) of global catch comes from TURFs and up to XX \% (0.55%) comes from cooperatives. In total, based on best available data to date,  we estimate that somewhere between XXX\% (9.8%) and XXX\% (11.1%) of global wild fish catch comes from all forms of RBFM combined. The What kinds of fisheries make up this RBFM catch? RBFM catches come from XX\% of the countries reporting  fish catches in 2016xx.

@@

library(dplyr)

edf_turf <- read_excel(here("data/EDF_TURF_data.xlsx"))%>%
  select("Country", "SciName","CatchShare") %>%
  filter(SciName != "NA")%>%
  mutate(turf = case_when(CatchShare == 1 ~ TRUE))

#DiscoverTURFs data
d_turfs <- read_excel(here("data/DiscoverTURFs_simple.xlsx")) %>%
  select("Country", "SciName", "CatchShare") %>%
  filter(SciName != "NA")%>%
    mutate(turf = case_when(CatchShare == 1 ~ TRUE))

projection_data_line70 <- readRDS(here("data","projection_data.rds")) %>% 
  filter(Scenario == "Historic")%>%
  separate(IdOrig, c("agency", "stock", "t0", "tf", "data_enterer"))%>%
  unite(assess_id_short, agency, stock, sep = "-") %>%
  group_by(assess_id_short) %>% 
  filter(Year == max(Year))%>%
  filter(Dbase != "SOFIA") %>%
  filter(IdLevel != "Neis")

countries_ITQ <- read_excel(here("data/Countries_ITQ.xlsx"))%>%
  group_by(assess_id_short) %>% 
  filter(year == max(year))

projection_corb <- left_join(projection_data_line70, countries_ITQ, by = c("assess_id_short"))

projection_corb_1 <- left_join(projection_corb, d_turfs, by = c("Country", "SciName"))

projection_corb_2 <- left_join(projection_corb_1, edf_turf, by = c("Country", "SciName"))

#####Turf math for upper and lower bounds.
###lower bounds just includes upside/corbette because those have fishery IDs associated with it so sure that those fisheies have turfs, the upper bound incorporates discover turfs and EDF data and links the country/scienfitic name of species to fisheries

turf_catch_lower <- projection_corb%>%
  filter(turf == "TRUE")
sum(turf_catch_lower$Catch)
#301,000 lower limit

turf_catch_upper <- projection_corb_2%>%
  filter(turf == "TRUE")
sum(turf_catch_upper$Catch)
#1,370,210 upper limit
#upper bound of total percent of global catch from turf:
1370210/80000000
#0.01712763

###cooperative estimates
coop_catch <- projection_corb%>%
  filter(coop == "TRUE")
sum(coop_catch$Catch)
#443231
#percent of global catch from coop:
443231/80000000
#0.005540387

###all rbfm estimate lower 
rbfm_catch_low <-projection_corb%>%
  filter(turf == "TRUE" | itq == "TRUE" | ivq == "TRUE" | iq == "TRUE" | coop == "TRUE")
sum(rbfm_catch_low$Catch)
#7863234
7863234/80000000
#0.09829043

###all rbfm estimate upper 
rbfm_catch_up <-projection_corb_2%>%
  filter(turf == "TRUE" | itq == "TRUE" | ivq == "TRUE" | iq == "TRUE" | coop == "TRUE")
sum(rbfm_catch_up$Catch)
#8932444
8932444/80000000
#0.1116556

#reported_2016 <- projection_corb_2 %>% filter(Year == "2016")

@@

What factors drive the adoption of various forms of RBFM? Anecdotal evidence suggests that large, industrial fisheries, often for mobile species are most likely to adopt some form of ITQ, while small-scale artisinal fisheries, often for benthic species, are most likely to adopt a TURF. To test this expectation, we built a dataset of XXX (94) ITQ fisheries and XXX (221) TURF fisheries, and matched them to the fishery classification in \cite{Upside}. This provided us with a dataset of XXX (121) fisheries with the country, species or genus name, and some other ecological and economic characteristics. Conditional on adopting some form of RBFM, our goal was to estimate the likelihood of adopting an ITQ versus a TURF. To do so, we ran a logistic regression with ITQ on the left ha

unique(turf_catch_upper$assess_id_short)
#27 fisheries with turfs

itqs <-projection_corb%>%
  filter(itq == "TRUE" | ivq == "TRUE" | iq == "TRUE")
unique(itqs$assess_id_short)
#94

turf_all_1 <- full_join(edf_turf, d_turfs, by= c("Country", "SciName"))
projection_turf <- projection_corb %>%
  filter(turf == "TRUE") %>%
  select("Country", "SciName")

turf_all <- full_join(projection_turf, turf_all_1, by= c("Country", "SciName"))%>%
  ungroup()%>%
  select(Country, SciName) %>%
  distinct()

#had information of 221 country/species combo for turf

@@

to all XX (5360) non-rbfm fisheries in the (however many fisheries are not catch shares in projection_data)

@@

projection_data_line107 <- readRDS(here("data","projection_data.rds")) %>% 
  filter(Scenario == "Historic")%>%
  group_by(IdOrig) %>% 
  filter(Year == max(Year))%>%
  select(IdOrig, CatchShare)%>%
  filter(CatchShare == "0") %>%
  distinct()
#5360 fisheries without catch share

@@

Using that approach, authors have estimated a benefit of global RBFM between \$XXX ( $31 billion) Billion and \$XXX ($53 billion) Billion per year. see Costello et al. 2016

1. Costello et al 2016 
2050

-	RBFM policy applied to just stocks of conservation concern = $31 billion in fisheries profit

-	RBFM policy applied to all stocks = $53 billion in fisheries profit

2. Original Sunken Billions (2009)

- “current annual net benefits from marine capture fisheries are tens of billions of U.S. dollars less than the potential benefits”
```{r}
include_graphics("sunkenbillionstable.png")
```

3. Updated Sunken Billions (2012)

-	economic losses of about $83 billion
